%LV begin insert content from D10
\title{Day 14: Project Presentations and Wrap-Up}

\section{Schedule}
\begin{itemize}
\item 9:00-9:45: Sharing Your Project
\item 9:45-10:15: Individual cheatsheet creation
\item 10:15-10:30 Break
\item 10:30-11:15 Collaborative mindmap creation

\item Free work time untl 12:10
\item 12:10-12:30 Wrap-up
\end{itemize}

\section{Lightning talks [90 minutes]}

Each team will have two minutes to present. If we can stay on schedule, we should have time for questions between talks and/or at the end.



\section{Break [15 minutes]}


\section{Synthesis, reflection, and feedback [60 min]}
\begin{enumerate}
\item \textbf{Final concept map for faces module [20 minutes]} In groups of 4-ish. You know the drill. Put a photo on Canvas. Once your group is done, check out other groups' maps.
\item \textbf{Individual reflection [20 minutes]} To start this process, we'd like to get some of your individual reflections on these questions.  Please complete the canvas survey on Module 2.
\item \textbf{Partner feedback form [5 minutes]} It is important to give people direct feedback.  Complete the provided partner feedback sheet, exchange it with your partner, and discuss. (If you didn't do the project, this would be a good time to check in the with teaching team about your progress.)
\item \textbf{Group retrospective [15 minutes]} Join with another table to make a group of 6-8 people, and run a group retrospective focused on what all of us (students, instructors and ninjas), as a class, have done for the last few weeks.  Spend a few minutes individually generating $+$ and $\Delta$ post-its that you'd like to share.  Narrow your feedback it to about 4-8 post-its total -- choose the important stuff! Then share with the group, organize your feedback on the board, and identify a couple of key messages to report out to the class and/or teaching team. Include a photo of your group's board in your day submission to Canvas.

\end{enumerate}

\section{Work time [?? minutes]}

First, sign up for  Module 3 NINJA check-ins \href{https://docs.google.com/spreadsheets/d/1vkMsrGzSqdBcEwj7fxO2WKl0Nr5Fm_EwxTAZLefAkxc/edit#gid=1131009963}{here}. Try to choose a time with a new NINJA, if you can.


Then, if you have time, either:

\begin{enumerate}
\item [A.] Go through the \href{https://docs.google.com/document/d/1zXXR0NJCqJshj_Dj2zSO54iw1vb4ec5UsukpeGv5jE8/}{Meeto your Neato} assignment (due Monday) and get set up as much as you can.
\item [B.] Work on your technical reports, drafts of which are due tomorrow!
\item [C.] If your draft technical report is already done, find an instructor to get early feedback.
\item [D.] Work on catching up on past assignments if you need to.
\end{enumerate}

\section{Wrap-up [20 minutes]}
%LV end insert content from D10